TITLE

### 2021-09-21 Appendix_test.Rmd
Add new appendix and revisions/reply to editor

---
title             : |
    | Does online masked priming pass the test?
    | The effects of prime exposure duration on masked identity priming
shorttitle        : "Online masked priming"
author:

  - name          : "Bernhard Angele"
    affiliation   : "1,2"
    corresponding : yes    # Define only one corresponding author
    address       : "Department of Psychology, Faculty of Science and Technology, Talbot Campus, Fern Barrow, Poole BH12 5BB, UK"
    email         : "bangele@bournemouth.ac.uk"
  - name          : "Ana Baciero"
    affiliation   : "2,3"
  - name          : "Pablo Gómez"
    affiliation   : "4"
  - name          : "Manuel Perea"
    affiliation   : "2,5"
affiliation:

  - id            : "1"
    institution   : "Bournemouth University, Bournemouth, UK"
  - id            : "2"
    institution   : "Universidad Antonio de Nebrija, Madrid, Spain"
  - id            : "3"
    institution   : "DePaul University, Chicago, USA"
  - id            : "4"
    institution   : "California State University San Bernardino, Palm Desert Campus, USA"
  - id            : "5"
    institution   : "Universitat de València, Valencia, Spain"
note: "\\clearpage"
authornote: |

  We thank Grace Backhouse for her help with data collection. The data for Experiment 1 were collected as part of her MSc dissertation.

  This research has been partly supported by Grant PSI2017-86210-P from  the  Spanish Ministry of Science and Innovation (MP), Grant PGC2018-097145-B-I00 from the Spanish State Research Agency (AB), and internal funding from Bournemouth University (BA).

  This study was pre-registered on the Open Science Framework (OSF) prior to data collection. The registration form for Experiment 1, together with the materials, data files, and R scripts can be found at [https://osf.io/57rzq/]. The registration form for Experiment 2 can be found at [https://osf.io/957s8].
abstract: |
  Masked priming is one of the most important paradigms in the study of visual word recognition, but it is usually thought to require a laboratory setup with a known monitor and keyboard. To test if this technique can be safely used in an online setting, we conducted two online masked priming lexical decision task experiments using PsychoPy/PsychoJS [@peirce2019]. Importantly, we also tested the role of prime exposure duration (33.3 vs. 50 ms in Experiment 1 and 16.7 vs. 33.3 ms in Experiment 2), thus allowing us to examine both across conditions and within-conditions effects. We found that our online data are indeed very similar to the masked priming data previously reported in the masked priming literature. Additionally, we found a clear effect of prime duration, with the priming effect (measured in terms of response time and accuracy) being stronger at 50 ms than 33.3 ms and no priming effect at 16.7 ms prime duration. From these results, we can conclude that modern online browser-based experimental psychophysics packages (e.g., PsychoPy) can present stimuli and collect responses on standard end user devices with enough precision. These findings provide us with confidence that masked priming can be used online, thus allowing us not only to run less time-consuming experiments, but also to reach populations that are difficult to test in a laboratory.

keywords          : "Masked priming, Lexical decision task, Online experiments, PsychoPy, Prime duration"
wordcount         : "7916"
floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no
csl               : "apa.csl"
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
bibliography: ["references.bib", "r-references.bib"]
editor_options: 
  chunk_output_type: console
---

@@

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
knitr::opts_chunk$set(error = TRUE)
```

Masked priming [@forsterRepetitionPrimingFrequency1984] is one of the most important techniques to study the effects of orthography, phonology, morphology, and meaning in visual word recognition [see @forster1998pros; @grainger2008cracking, for reviews]. Priming refers to the influence of a prime stimulus (e.g., *nurse*, *horse*) on a subsequently presented stimulus that the participant has to respond to (e.g., "is *DOCTOR* a word?"). Priming effects are measured as the difference in a dependent variable (e.g., response time [RT]) between two conditions (e.g., unrelated: *horse-DOCTOR*; related: *nurse-DOCTOR*). In masked priming, the prime stimulus is presented very briefly (for less than 60 ms) and is itself preceded by a pattern mask (e.g., \#\#\#\#\#) for a much longer duration (typically 500 ms). The rationale of the procedure is to make participants unaware of the identity of the masked prime [@forsterRepetitionPrimingFrequency1984; @forster1998pros], thus minimizing the role of participants' strategies. Indeed, masked priming experiments do not show the strategical effects that occur with visible, unmasked primes [e.g., @grossi2006relatedness; @perea2002does].

The masked priming paradigm has been used in a large number of studies over the last decades. For instance, a search of the expression "masked priming" in Google Scholar in May 2021 produced more than 10,000 hits. Nearly all masked priming experiments have been run in a laboratory setting, often using the DMDX software developed by @forsterDMDXWindowsDisplay2003. The issue we examine in the present paper is whether masked priming experiments can be conducted in an online setting without significant changes in the pattern of results. Even before the exceptional situation due to the COVID pandemic in 2020-21, in which many labs around the world were closed (or reduced their activity) for many months, online data collection has shown its many advantages: 1) easy access to a much more diverse population than that accessible at the typical university research laboratory; 2) independence from laboratory space constraints, and, often, lower costs as participants only need to be compensated for their time on the experiment; 3) no time spent commuting, waiting for the experiment to start, etc. Indeed, researchers in decision making and economics have been using online paradigms for several decades now [e.g., @birnbaumPsychologicalExperimentsInternet2000a; @paolacciRunningExperimentsAmazon2010].

Cognitive psychologists have been much slower that their behavioral economics colleagues in taking up online paradigms [see @brysbaert2016impact; @cai2017accent; @Dufau_2011; @eerland2013influence; @rodd2016impact, for some exceptions], often due to concerns about the validity of the results. Such concerns are not limited to cognitive studies [see @austSeriousnessChecksAre2013], but they are exacerbated by the reliance on precise presentation times in cognitive psychology. Furthermore, these concerns are even more central in the masked priming technique, where it is critical for the onset of the mask, the prime, and the target to occur at the nominal times. For instance, presenting the masked prime for longer than intended (e.g., 83 ms or longer instead of the nominal 50 ms) could counteract the effect of the mask, making the prime consciously visible to the participant and possibly altering the processes of interest [see @Zimmerman_2012].

There have been attempts to address these concerns. @witzelTestingViabilityWebDMDX2013 developed a Web version of DMDX (webDMDX) showing promising results in a trial experiment. However, webDMDX is a self-contained Windows executable file that participants have to download and run rather than a "true" online programming script that could be run inside of a browser. A downside of this format is that participants often are (and should be) understandably skeptical about downloading and running executable files from the Internet. Additionally, many participants may not have access to a Windows PC, or may be discouraged from participating by the extra work it takes to deploy the experiment on their computer. As a consequence, the use of webDMDX in masked priming experiments has been rather limited so far [see @Alluhaybi_2020; @Dubey_2018, for exceptions].

Fortunately, in recent years, there have been significant improvements in how content can be presented on the World Wide Web. Most notably, the HTML5 standard now makes it possible to use Javascript in order to draw stimuli interactively and monitor participant responses with remarkable flexibility inside the browser. Participants do not have to install any software, and the HTML5 standard is supported by a wide variety of devices, including mobile phones and tablets [@reimersPresentationResponseTiming2015]. There is now a variety of software packages taking advantage of the new capabilities to present experimental stimuli and collect data, both commercial, such as Gorilla [@anwyl-irvine2020] or Testable [@rezlescu2020], and open-source such as jsPsych [@deleeuw2015] or PsychoJS [the Javascript version of PsychoPy 3, @peirce2019]. In addition, online setups allow researchers to target any individual with an internet connection as a participant, from very different countries and backgrounds. Indeed, various on-line platforms (e.g., Prolific, Amazon Mechanical Turk) offer the possibility of recruiting participants for on-line experiments based on various specific characteristics set up by the experimenters regardless of their location (e.g., native French speakers, not older than 30 years old, not currently in college).

Of course, despite the technological advances, many cognitive psychologists still have concerns about timing and measurement precision. While Javascript-based experiments run on the participants' devices and thereby avoid any lag due to connection issues (e.g., to avoid delays, all stimuli are usually downloaded before the start of the experiment), experimenters have little control over which devices the experiment is run on beyond the option of explicitly preventing the experiment to run on specific device types such as mobile devices. Moreover, experimenters have no control at all over what other applications are running on the device, screen size and resolution, viewing distance, properties of the keyboard/touchscreen, etc., as all of these are determined by the device or the participants' preferences. As @reimersPresentationResponseTiming2015 pointed out, there are two ways of testing whether timing and response issues are problematic: (1) comparing a Web-based experiment directly with an established lab-based version by measuring presentation timings (using a photodiode) and response timings on various device configurations and (2) attempting to replicate existing lab-based findings using a Web-based paradigm. If the results of the Web-based study are comparable to previous lab-based results, this suggests that, whatever the deviations in stimulus and response timing are, they are not severe enough to affect the overall findings in the paradigm in question.

The first approach has the advantage that differences in presentation timings can be objectively recorded and evaluated. A very thorough recent example of this approach is the "timing mega-study" by @bridges2020, who compared the timing in experiments run in lab-based setups with the timing in online packages run in different browsers. The very similar study by @anwyl-irvine2020a compares only online packages and browsers with regard to timing. Overall, @bridges2020 found that online packages were capable of presenting visual stimuli with reasonable precision, although the lab-based packages were slightly better in this regard. This first approach is important in order to establish that a certain level of precision and accuracy can be achieved at all. If this is not possible, there is no point in moving forward to the second approach and replicating specific paradigms. However, it is of course impossible to test every possible device and configuration that participants might use. On the other hand, some of the differences in precision and accuracy between setups that can be observed using a photodiode may be too small to have an influence on actual participant performance. Therefore, we consider replication of previous key lab-based effects a more important test of online paradigms than photodiode measurements. Based on the results by @bridges2020 and @anwyl-irvine2020a, modern Javascript-based stimulus presentation systems are capable of sufficiently fast and precise stimulus presentation. To establish whether masked priming studies can be successfully run online, the next step is now to follow the second approach and implement the masked priming paradigm online and test whether results obtained via in-lab studies can be replicated, which is at the heart of the present study. Importantly, we will do so using PsychoPy/PsychoJS [@peirce2019] as it showed high precision and accuracy across the great majority of platforms [@bridges2020; @anwyl-irvine2020a], in addition to being an open-source software.

Specifically, in this study we were interested in whether we could replicate and extend a key phenomenon in laboratory masked priming lexical decision using an online setup: masked identity priming is commonly described as a savings effect. As first suggested by @forster1998pros, for a masked identity prime, "the lexical entry is already in the process of being opened, and hence the evaluation of this entry begins sooner", whereas for an unrelated prime, "the entry for the target word would be closed down (since it fails to match the prime), and no savings would occur" (p. 213). Thus, according to the savings account, a target word like *DOCTOR* would enjoy an encoding advantage when preceded by an identity prime such as *doctor* than when preceded by an unrelated prime such as *pencil* (i.e., a head-start). One implication of such benefit is that the RT distributions of the unrelated and identity pairs should reflect a shift rather than a change in shape. Furthermore, this shift should be approximately similar in magnitude to the prime-target stimulus-onset asynchrony (SOA). Empirical evidence supporting this view has been obtained in several studies not only with skilled adults but also with developing readers [e.g., @gomezDiffusionModelAccount2013; @Gomez_2020; @Taikh_2020; @Yang_2021]. @gomezDiffusionModelAccount2013 proposed an implementation of this hypothesis within the diffusion model [@ratcliff2004]. This implementation proposes that, when making a two-choice decision, the resultant RT can be explained as the sum of non-decision parameters, which are the encoding time and response execution ($T_{er}$) and decision parameters, which refer to the process of accumulation of information until a decision criteria is reached. Importantly, in the decision process, the information gathered from the stimulus can vary in noise, depending on its quality, which modifies the rate at which information is accumulated (i.e., the *drift rate*). With regards to RTs from masked priming tasks, @gomezDiffusionModelAccount2013 found that the difference between identity and unrelated conditions could be accounted for by a change in the $T_{er}$ parameter, while there were no differences across conditions in the parameter that corresponds to the quality of evidence gathered (i.e., drift rate)--note that changes in drift rate would necessarily produce a more skewed RT distribution in the slower, unrelated condition.

Critically, the above pattern is specific to *masked* priming. In fact, the most common pattern of results in latency based tasks is that conditions that produce longer latencies will also produce larger variance. This pattern is evident in priming as well---when primes are visible (i.e., unmasked priming), identity priming effects are stronger in the upper quantiles of the RT distribution than in the lower quantiles [i.e., a change in shape rather than a shift in RT distributions; see @gomezDiffusionModelAccount2013]. Fits from the diffusion model show that this result corresponds to changes in both the $T_{er}$ parameter and the drift rate. Hence, when the prime is visible, it does influence the quality of the information accumulated of the target word, unlike in masked priming. Clearly, this dissociation between masked and unmasked priming reflects qualitative differences in the way primes affect the processing of the target: purely encoding in masked priming (with an expected effect close to the prime duration) vs. both encoding and information quality in unmasked priming.

In the present paper, we took advantage of the above marker to examine whether online masked priming studies follow the same pattern as in-lab masked priming studies. Specifically, we manipulated prime exposure duration in identity vs. unrelated primes: 33.3 vs. 50 ms in Experiment 1, and 16.7 vs. 33.3 ms in Experiment 2---note that targets were presented immediately after the primes (i.e., prime exposure duration was equal to the prime-target SOA). The rationale of Experiment 1 is that if the actual exposure duration of the primes is the nominal exposure duration, then we would expect the typical shift between the identity and unrelated response time distributions, which according to the savings hypothesis [@forster1998pros] would be greater for 50 ms than for 33.3 ms exposure duration (i.e., the head-start would be greater for 50 ms identity primes than for 33.3 ms identity primes). This outcome would indicate that the on-line masked priming studies reproduce a characteristic signature of laboratory masked priming studies. Alternatively, if the online presentation conditions lead to a greater actual exposure duration on the participant's device compared to that specified in our experiment (e.g.,if prime durations were, on average, 20 ms longer than intended), then the 50 ms primes may no longer be adequately masked, but may rather be consciously perceived. If this is the case, the prime could affect not only the encoding, but also core decision processes (i.e., the drift rate), which would be reflected as a stronger priming effect in the higher quantiles of the distribution (i.e., the two RT distributions would have a different shape). In this scenario, one should be very cautious when running online masked priming experiments---at least with the typical software and hardware currently available.

To further constrain the research questions, Experiment 2 was designed to be analogous to Experiment 1, except for replacing the 50 ms prime exposure duration with a very short prime exposure duration, namely, 16.7 ms. Similar extremely brief prime exposure durations have shown very weak masked priming lexical decision experiments in a laboratory setting: less than 5 ms for a prime duration of 14 ms [@ziegler2000] and less than 9 ms for a prime duration of 20 ms [@Tzur_2007]---in the @Tzur_2007 experiment, this difference increased to 16.7 ms when using a very high level of contrast in the computer screen. Thus, if the size of the priming effect is roughly similar to the prime exposure duration, we would expect a much larger priming effect at the 33.3 ms prime exposure duration than at the 16.7 ms prime exposure duration. If we do observe a large effect at the 16.7 ms prime exposure duration (e.g., above 20-25 ms), this would suggest, again, that there is a qualitative difference between using the masked priming technique in the laboratory and in online experiments. Keep in mind that we are using a software that has very good control over the exposure duration [@bridges2020; @anwyl-irvine2020a]. Figure \ref{fig:hyp-mod-dat-figure} displays the link between all these verbal hypotheses, the process model that corresponds to each of them, and their predictions in the RT distributions.

All in all, replicating the results observed in controlled laboratory masked priming studies would be an important step in establishing the validity of online masked priming tasks, even in a scenario with no control over many variables (e.g., the devices in which the experiment is run, the level of contrast of the computer screen, or the additional applications that are running in the background of said device), as has been done with other paradigms that also require a high control of the presentation timing and location of the stimuli [e.g., @Parker_2021]. This would open the possibility for masked priming researchers to collect large samples from diverse populations via online crowdsourcing methods, as has been done previously using other paradigms [e.g., @aguasvivas2020; @brysbaert2021a; @mandera2020; @brysbaert2016impact]. So far, this has only been possible in many-labs studies depending on the collaboration of many researchers in different countries [e.g., @adelman2014]. Importantly, the present experiments go beyond an online replication: they allow us to test masked priming effects not only across conditions (identity vs. unrelated) but also within conditions (the effect of identity primes, or unrelated primes, across two prime durations) [see @Jacobs_1995; @ziegler2000, for advantages of within-condition comparisons]. Thus, the within-subject manipulation of prime exposure duration in both experiments allows us to obtain a comprehensive picture of masked identity priming effects--also including the examination of potential inhibitory effects of unrelated primes.

\newpage

```{r hyp-mod-dat-figure, echo=FALSE, as.is = TRUE, warning=FALSE, message=FALSE, fig.height=3.4, fig.cap = "Mapping from hypotheses to data via evidence accumulation models. The bottom part of each graph represents sample paths, and the top part shows the cumulative density functions.  From left to right: *Null Effect* - there are no differences between the conditions; *Effect in Encoding* - the difference between the conditions is on the encoding time, hence the RT distributions are shifted; and *Effect of Encoding and Evidence Accumulation* - the differences between the conditions is on the rate of evidence accumulation, thus increasing the effect size for longer RTs (right-tail of the distribution)."}

suppressMessages(library("tidyverse"))

#pdf("Fig1_ModelsToData.pdf", height = 3)

par(mfrow=c(1,3))
tdp   <- .386
sd.dp <- .100

nsims <- 1000000

slope.gray  <- 0.00001
slope.blue   <- rnorm(nsims, 3, .6)
slope.yellow <- rnorm(nsims, 3, .6)

sim.trials.blue<-matrix(ncol=4, nrow=nsims)
sim.trials.yellow<-matrix(ncol=4, nrow=nsims)

sim.trials.blue[,1]    <-  rep(tdp+.05, nsims)
sim.trials.yellow[,1]  <-  rep(tdp+.05, nsims)

sim.trials.blue[ ,2]   <- (slope.gray * sim.trials.blue[,1])
sim.trials.yellow[ ,2] <- (slope.gray * sim.trials.yellow[,1])

sim.trials.blue[ ,3]   <- (sim.trials.blue[,2] - slope.blue * sim.trials.blue[,1])
sim.trials.yellow[ ,3] <- (sim.trials.yellow[,2] - slope.yellow * sim.trials.yellow[,1])

sim.trials.blue[ ,4]   <- (1 - sim.trials.blue[,3]) / slope.blue
sim.trials.yellow[ ,4] <- (1 - sim.trials.yellow[,3]) / slope.yellow

nlines<- 1000
sample.blue <- sample(1:nsims,nlines)
sample.yellow <- sample(1:nsims,nlines)

plot(c(rep(0,nlines),sim.trials.blue[sample.blue,1]), c(rep(.02,nlines),.02+sim.trials.blue[sample.blue,2]), type="l", col=alpha("dodgerblue2",alpha = .5), ylim = c(0,1.7), xlim=c(0,1.2), bty="n",yaxt='n', xlab="RT (sec)", ylab="", lwd=3)

lines(c(rep(0,nlines),sim.trials.yellow[sample.yellow,1]), c(rep(0,nlines),sim.trials.yellow[sample.yellow,2]), type="l", col=alpha("deeppink2",alpha = .5), lwd=3)

for(i in 1:nlines){
    lines(c(sim.trials.blue[sample.blue[i],1], sim.trials.blue[sample.blue[i],4]), c(sim.trials.blue[sample.blue[i],2], 1), col=alpha("dodgerblue2",alpha = .005),lwd=2)

    lines(c(sim.trials.yellow[sample.yellow[i],1], sim.trials.yellow[sample.yellow[i],4]), c(sim.trials.yellow[sample.yellow[i],2], 1), col=alpha("deeppink2",alpha = .005), lwd=2)
}

lines(density(sim.trials.yellow[,4], from=.4,to=1.5)$x,(density(sim.trials.yellow[,4],, from=.4,to=1.5)$y)/20+1, col=alpha("deeppink2",alpha = 1), lwd=1)
lines(density(sim.trials.blue[,4], from=.4,to=1.5)$x,(density(sim.trials.blue[,4], from=.4,to=1.5)$y)/20+1, col=alpha("dodgerblue2",alpha = .3),lwd=4)

lines(c(0, 1), c(1,1), lwd=1.5)

lines(c(tdp, tdp), c(0,1), lty=2)
lines(c(tdp+.05, tdp+.05), c(0,1), lty=2)

text("Evidence\naccumulation",x = .88, y=.4,cex=.8)
text("Encoding",x = .2, y=.4,cex=.8)
title("Null Effect")

###########################################################################

tdp   <- .386
sd.dp <- .100

nsims <- 1000000

slope.gray  <- 0.00001
slope.blue   <- rnorm(nsims, 3, .6)
slope.yellow <- rnorm(nsims, 3, .6)

sim.trials.blue<-matrix(ncol=4, nrow=nsims)
sim.trials.yellow<-matrix(ncol=4, nrow=nsims)

sim.trials.blue[,1]    <-  rep(tdp, nsims)
sim.trials.yellow[,1]  <-  rep(tdp+.05, nsims)

sim.trials.blue[ ,2]   <- (slope.gray * sim.trials.blue[,1])
sim.trials.yellow[ ,2] <- (slope.gray * sim.trials.yellow[,1])

sim.trials.blue[ ,3]   <- (sim.trials.blue[,2] - slope.blue * sim.trials.blue[,1])
sim.trials.yellow[ ,3] <- (sim.trials.yellow[,2] - slope.yellow * sim.trials.yellow[,1])

sim.trials.blue[ ,4]   <- (1 - sim.trials.blue[,3]) / slope.blue
sim.trials.yellow[ ,4] <- (1 - sim.trials.yellow[,3]) / slope.yellow

nlines<- 1000
sample.blue <- sample(1:nsims,nlines)
sample.yellow <- sample(1:nsims,nlines)

plot(c(rep(0,nlines),sim.trials.blue[sample.blue,1]), c(rep(.02,nlines),.02+sim.trials.blue[sample.blue,2]), type="l", col=alpha("dodgerblue2",alpha = .5), ylim = c(0,1.7), xlim=c(0,1.2), bty="n",yaxt='n', xlab="RT (sec)", ylab="", lwd=3)

lines(c(rep(0,nlines),sim.trials.yellow[sample.yellow,1]), c(rep(0,nlines),sim.trials.yellow[sample.yellow,2]), type="l", col=alpha("deeppink2",alpha = .5), lwd=3)

for(i in 1:nlines){
    lines(c(sim.trials.blue[sample.blue[i],1], sim.trials.blue[sample.blue[i],4]), c(sim.trials.blue[sample.blue[i],2], 1), col=alpha("dodgerblue2",alpha = .006),lwd=2)

    lines(c(sim.trials.yellow[sample.yellow[i],1], sim.trials.yellow[sample.yellow[i],4]), c(sim.trials.yellow[sample.yellow[i],2], 1), col=alpha("deeppink2",alpha = .006), lwd=2)

}

lines(density(sim.trials.yellow[,4], from=.4,to=1.5)$x,(density(sim.trials.yellow[,4],, from=.4,to=1.5)$y)/20+1, col=alpha("deeppink2",alpha = .7), lwd=2)
lines(density(sim.trials.blue[,4], from=.4,to=1.5)$x,(density(sim.trials.blue[,4], from=.4,to=1.5)$y)/20+1, col=alpha("dodgerblue2",alpha = .7),lwd=2)

lines(c(0, 1), c(1,1), lwd=1.5)

lines(c(tdp, tdp), c(0,1), lty=2)
lines(c(tdp+.05, tdp+.05), c(0,1), lty=2)

text("Evidence\naccumulation",x = .88, y=.4,cex=.8)
text("Encoding",x = .2, y=.4,cex=.8)

title("Effect in\nEncoding")

#############################################################################

tdp   <- .386
sd.dp <- .100

nsims <- 1000000

slope.gray  <- 0.00001 
slope.blue   <- rnorm(nsims, 3.8, .6)
slope.yellow <- rnorm(nsims, 3, .6)

sim.trials.blue<-matrix(ncol=4, nrow=nsims)
sim.trials.yellow<-matrix(ncol=4, nrow=nsims)

sim.trials.blue[,1]    <-  rep(tdp, nsims)
sim.trials.yellow[,1]  <-  rep(tdp+.05, nsims)

sim.trials.blue[ ,2]   <- (slope.gray * sim.trials.blue[,1])
sim.trials.yellow[ ,2] <- (slope.gray * sim.trials.yellow[,1])

sim.trials.blue[ ,3]   <- (sim.trials.blue[,2] - slope.blue * sim.trials.blue[,1])
sim.trials.yellow[ ,3] <- (sim.trials.yellow[,2] - slope.yellow * sim.trials.yellow[,1])

sim.trials.blue[ ,4]   <- (1 - sim.trials.blue[,3]) / slope.blue
sim.trials.yellow[ ,4] <- (1 - sim.trials.yellow[,3]) / slope.yellow

nlines<- 1000

sample.blue <- sample(1:nsims,nlines)
sample.yellow <- sample(1:nsims,nlines)

plot(c(rep(0,nlines),sim.trials.blue[sample.blue,1]), c(rep(.02,nlines),.02+sim.trials.blue[sample.blue,2]), type="l", col=alpha("dodgerblue2",alpha = .5), ylim = c(0,1.7), xlim=c(0,1.2), bty="n",yaxt='n', xlab="RT (sec)", ylab="", lwd=3)

lines(c(rep(0,nlines),sim.trials.yellow[sample.yellow,1]), c(rep(0,nlines),sim.trials.yellow[sample.yellow,2]), type="l", col=alpha("deeppink2",alpha = .5), lwd=3)

for(i in 1:nlines){
    lines(c(sim.trials.blue[sample.blue[i],1], sim.trials.blue[sample.blue[i],4]), c(sim.trials.blue[sample.blue[i],2], 1), col=alpha("dodgerblue2",alpha = .006),lwd=2)

    lines(c(sim.trials.yellow[sample.yellow[i],1], sim.trials.yellow[sample.yellow[i],4]), c(sim.trials.yellow[sample.yellow[i],2], 1), col=alpha("deeppink2",alpha = .006), lwd=2)

}

lines(density(sim.trials.yellow[,4], from=.4,to=1.5)$x,(density(sim.trials.yellow[,4],, from=.4,to=1.5)$y)/20+1, col=alpha("deeppink2",alpha = .7), lwd=2)
lines(density(sim.trials.blue[,4], from=.4,to=1.5)$x,(density(sim.trials.blue[,4], from=.4,to=1.5)$y)/20+1, col=alpha("dodgerblue2",alpha = .7),lwd=2)

lines(c(0, 1), c(1,1), lwd=1.5)

lines(c(tdp, tdp), c(0,1), lty=2)
lines(c(tdp+.05, tdp+.05), c(0,1), lty=2)

text("Evidence\naccumulation",x = .88, y=.4,cex=.8)
text("Encoding",x = .2, y=.4,cex=.8)

title("Effect in Encoding and \nEvidence Accumulation")

#dev.off()

# Experiment 1

In the first experiment, we tested whether we could observe reliable effects of masked priming at prime durations of 33.3 ms and 50 ms (roughly corresponding to two and three frames, respectively, at a refresh rate of 60 Hz).

We report here how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the two experiments. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

## Method

```{r load_exp1_data, message=FALSE, echo=FALSE, warning=FALSE}

# Qualtrics survey data
exp1_all_participants <- read_csv("participant_data_exp1.csv")

# PsychoJS trial data
exp1 <- fs::dir_ls(path = "./data_exp1", glob = "*.csv") %>%
  map_dfr(read_csv, .id = "source", col_type = cols(
  .default = col_character(), rt = col_double(), corr = col_integer(), TrialID = col_integer())) %>% 
  filter(!is.na(TrialID) & TrialID < 1000) %>%
  select(source, participant, date, OS, frameRate, rt, corr, TrialID, StimulusType, Condition, PrimeDuration, Prime, Target)

exp1$device <- ifelse(exp1$OS %in% c("Linux armv7l", "Linux armv8l"), "android", "computer")

# only consider participants who have actually participated in the whole experiment
# also don't consider participants whose age is 99 (used for testing)
# one participant is duplicate. Remove the first occurrence (likely incompatible device)

exp1_actual_participants <- filter(exp1_all_participants, (PROLIFIC_PID %in% exp1$participant) & (nchar(PROLIFIC_PID) == 24) & !(PROLIFIC_PID %in% c("5fa3b4abbcfd0b6c243758bc")) & (`What is your age?` != 99) & (`Response ID` != "R_sU4qJ6UCe9jRKs9"))

# get participants with more than 80% correct responses

exp1_accuracy_by_participant <- exp1 %>% filter(participant %in% exp1_actual_participants$PROLIFIC_PID) %>% group_by(source) %>% summarise(acc = mean(corr == 1), N = n())

exp1_participants_to_include <- exp1_accuracy_by_participant %>% filter(N == 480 & acc >= .8)

exp1_participants_to_exclude <- exp1_accuracy_by_participant %>% filter(N != 480 | acc < .8)

#participant_list_for_prolific <- str_extract(string = participants_to_include$source, pattern = "[a-z0-9]{24}")

exp1_data_to_include <- exp1 %>% filter(source %in% exp1_participants_to_include$source & participant %in% exp1_actual_participants$PROLIFIC_PID) %>% mutate(StimulusType = StimulusType %>% factor(levels = c("NW", "WORD"), labels = c("Nonword", "Word")), Relatedness = Condition %>% factor(levels = c("ID", "UN"), labels = c("Identity", "Unrelated")), PrimeDuration = PrimeDuration %>% factor(levels = c(33, 50), labels = c("33.3 ms", "50 ms")), rt = rt * 1000)

# filter out any of the participants we eliminated from the Qualtrics file as well

exp1_actual_participants <- filter(exp1_actual_participants, (PROLIFIC_PID %in% exp1_data_to_include$participant))

exp1_participant_ages <- exp1_actual_participants$`What is your age?`

exp1_male_participants_nr <- filter(exp1_actual_participants, `What is your gender (optional)?` == "Male") %>% nrow

exp1_female_participants_nr <- filter(exp1_actual_participants, `What is your gender (optional)?` == "Female") %>% nrow

# there aren't actually any of these, but will leave the code in just in case
exp1_other_participants_nr <- filter(exp1_actual_participants, `What is your gender (optional)?` == "Other") %>% nrow

exp1_NA_participants_nr <- filter(exp1_actual_participants, is.na(`What is your gender (optional)?`)) %>% nrow

exp1_participant_coords <- exp1_actual_participants %>% select("Location Longitude", "Location Latitude")

#coordinates(exp1_actual_participants) <- c("Location Longitude","Location Latitude")

# from https://stackoverflow.com/questions/14334970/convert-latitude-and-longitude-coordinates-to-country-name-in-r

# The single argument to this function, points, is a data.frame in which:
#   - column 1 contains the longitude in degrees
#   - column 2 contains the latitude in degrees
coords2country = function(points)
{  
  countriesSP <- getMap(resolution='low')
  #countriesSP <- getMap(resolution='high') #you could use high res map from rworldxtra if you were concerned about detail

  # convert our list of points to a SpatialPoints object

  # pointsSP = SpatialPoints(points, proj4string=CRS(" +proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0"))

  #setting CRS directly to that from rworldmap
  pointsSP = SpatialPoints(points, proj4string=CRS(proj4string(countriesSP)))

  # use 'over' to get indices of the Polygons object containing each point 
  indices = over(pointsSP, countriesSP)

  # return the ADMIN names of each country
  indices$ADMIN  
  #indices$ISO3 # returns the ISO3 code 
  #indices$continent   # returns the continent (6 continent model)
  #indices$REGION   # returns the continent (7 continent model)
}

participant_coords <- exp1_actual_participants %>% select("Location Longitude", "Location Latitude")

exp1_actual_participants$country <- coords2country(participant_coords)

exp1_participant_countries <- table(exp1_actual_participants$country %>% as.character) #make into a character vector since the original output is a factor with as many levels as there are countries
```

The pre-registration form for Experiment 1, together with the materials, data files, and R scripts can be found at <https://osf.io/57rzq/>.

### Participants

Participants were recruited through Prolific [www.prolific.co, -@prolific2021]. The experiment was accessed by `r exp1_all_participants$PROLIFIC_PID %>% unique %>% length` participants. Out of these, `r exp1_accuracy_by_participant %>% nrow` provided experimental data. `r exp1_participants_to_exclude %>% filter(N != 480) %>% nrow %>% numbers_to_words() %>% str_to_sentence()` of these participants did not complete the experiment. A further `r exp1_participants_to_exclude %>% filter(N == 480 & acc < .8) %>% nrow %>% ifelse(. < 11, numbers_to_words(.), .)` were excluded because of low accuracy (less than .8, which was the pre-registered criterion). In the end, we analyzed the data from `r exp1_actual_participants$PROLIFIC_PID %>% unique %>% length` participants (`r exp1_female_participants_nr` female), aged from `r exp1_participant_ages %>% min` to `r exp1_participant_ages %>% max` (mean age: `r exp1_participant_ages %>% mean %>% round(2)`). All these participants indicated that English was their first language in the Prolific screening questions. Based on their IP addresses, `r exp1_participant_countries["United Kingdom"] %>% ifelse(. < 11, numbers_to_words(.), .)` participants were based in the UK, `r exp1_participant_countries["United States of America"] %>% ifelse(. < 11, numbers_to_words(.), .)` were based in the US, `r exp1_participant_countries["Canada"] %>% ifelse(. < 11, numbers_to_words(.), .)` participants were based in Canada, and `r exp1_participant_countries["Ireland"] %>% ifelse(. < 11, numbers_to_words(.), .)` participants were based in Ireland. Two participants could not be localized in this way. All participants were naïve to the purpose of the experiment and received £1.25 for their participation (corresponding to £5/hour). Participants could use either a desktop/laptop computer or a mobile device. Because of a technical display issue with PsychoJS and the Safari browser, participants who tried to access the experiment using that browser, including all participants on iOS devices, were advised to change browser or device and restart the experiment.

***Rationale for sample size and stopping rule.*** @brysbaert2018 recommended that masked priming experiments should have at least 1,600 observations per condition. In order to account for potentially smaller effect sizes in an online experiment, we set a target of 3,000 observations per condition (12,000 observations total), which corresponds to a minimum of 50 participants given that there were 240 word stimuli in the experiment (see below). Our stopping rule was to keep collecting data until 3,000 valid observations were found. This goal was met and exceeded in our initial data collection with a budget of £200.

### Materials

>>>>

---
appendix: 
  - "./Appendix_A.Rmd"
  - "./Appendix_B.Rmd"
output: papaja::apa6_pdf
---

This is the manuscript.

@@
